---
title: "mini_report_joseph_kilgallen"
author: "Joseph Kilgallen"
date: "11/19/2020"
output: html_document
---
# Exploratory Report: Juvenile Snowshoe Hares of Bonanza Creek Experimental Forest

Create your organized mini exploratory report in an R Markdown document. The knitted document should include only your final text and figures (code, exploratory analyses, messages and warnings should not appear in the knitted document). 

```{r setup, include=FALSE, message = FALSE, warning = FALSE, echo =FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Loading packages
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(ggbeeswarm)
library(effsize)
library(broom)
library(ggpubr)
library(kableExtra)
# Reading in the data
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv"))
# Creating a subset of just juvenile hares to do some more descriptive stats on
hares_juvenile <- bonanza_hares %>% 
  select(c("date", "sex", "weight", "hindft", "b_key", "age", "grid", "trap")) %>% 
  filter(age == "j") %>% 
  filter(!is.na(weight)) %>% 
  filter(!is.na(hindft)) %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) 
```

## 1. Introduction: 

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last.

## 2. Data and analyses: 

A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.



## 3. Exploratory findings: 
```{r,echo =FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 1. Total Juvenile Snowshoe Hare Trappings by Year- 1999-2012- sampled from the Bonanza Creek Experimental Forest, years 2002 and 2009 exempted due to 0 recored observations"}

# Doing some explorative analyses to understand the data and distributions
weight_histogram <- ggplot(hares_juvenile, aes(x = weight, color = sex)) +
  geom_histogram() +
  facet_wrap(~sex)
hindft_histogram <- ggplot(hares_juvenile, aes(x = hindft, color = sex)) +
  geom_histogram() +
  facet_wrap(~sex)
# now comparing the two groups means for hndft and weight
mean_comp_wthndft <- ggplot(hares_juvenile, aes(x = weight, y = hindft, color = sex)) +
  geom_point() +
  facet_wrap(~sex)
# Appears that the data by sex is somewhat normally distributed, with females reporting generally greater weights, perhaps this is being confounded by the samples for each sex?

# Annual juvenile hare trap counts. Count the total number of juvenile (age == "j") hare trappings during each year of the study (include all grid sites), and create a finalized data visualization of the counts by year. Include a figure caption below the figure in your report. 


traps_per_year <- hares_juvenile %>% 
  count(trap_new = round(year))

traps_summary <- traps_per_year %>% 
  summarize(mean = mean(n),
            sd = sd(n),
            median = median(n))
  
 
trap_plot <- ggplot(data = traps_per_year, aes(x = factor(trap_new), y= n)) +
  geom_col(color = "red", fill = "blue") +
  labs(x = "Year", y = "Total Juvenile Hares Trapped", title = "Juvenile Hares Trapped by Year from Bonanza Creek LTER") +
  theme_light()

trap_plot
```
Assessing the results of our exploratory visualization of juvenile hares trapped by year in the Bonanza Creek LTER, we find that the largest number of hares trapped in one of the recorded years (1999-2012) was 79 hares; successfully trapped in 1999. For the years 2002 and 2009 we have no recorded hares trapped, potentially indicating a lack of trapping activities in these year as all other years were at least successful in trapping one hare at a minimum (see 2010). While actual mean annual number of hares trapped was found to be roughly 21 hares per year and median annual hare number trapped was 13.5 hares, in general there appears to be a pattern of declining hare capture across the years. I posit that unequal effort (in terms of time, traps used, people engaged in this work) was used during the recorded years of trapping thus leading to the high variation in trap success across years. I would suggest creating a measure of trapping effort so we could calculate the level of trapping success across years as a function of effort employed.

### A. Visualize juvenile hare weights. 
```{r, echo =FALSE, message = FALSE, warning = FALSE, fig.cap= "Figure 1. Comparison of Juvenile Snowshoe Hare Weights by Sex and site of data collection. Data from the Bonanza Experimental Forest 1998-2012"}
# Create a finalized data visualization (or visualizations) in which you compare **juvenile hare weights by sex and site (grid)** (ignore other variables). You should include groups where these variables are not recorded (`NA`). Include finalized figure captions below your visualization(s).
juvenile_weight_viz <- bonanza_hares %>% 
  select(c("date", "sex", "weight", "hindft", "grid", "age")) %>% 
  filter(age == "j") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(sex = case_when(
    sex == "f" ~"Female",
    sex == "m" ~"Male",
  )) %>% 
  mutate(grid_full = case_when(
    grid == "bonbs" ~"Bonanza Black Spruce",
    grid == "bonmat" ~"Bonanza Mature",
    grid == "bonrip" ~"Bonanza Riparian",
  ))

# making means for the beeswarm
mean_weight_male <- hares_juvenile %>% 
  filter(sex == "m") %>% 
  filter(!is.na(weight)) %>% 
  pull(weight)
mean_weight_female <- hares_juvenile %>% 
  filter(sex == "f") %>% 
  filter(!is.na(weight)) %>% 
  pull(weight)

# Summary values made to make the beeswarm
hares_summary <- hares_juvenile %>% 
  group_by(sex) %>% 
  summarize(mean_wt = mean(weight, na.rm = TRUE), median_wt = median(weight, na.rm = TRUE), sd_wt= sd(weight, na.rm = TRUE), 
            mean_hft = mean(hindft, na.rm = TRUE), median_hft = median(hindft, na.rm = TRUE), sd_hft= sd(hindft, na.rm = TRUE), 
            sample_size= n())

# making chart
weightplot <- ggplot() + 
  geom_beeswarm(data = juvenile_weight_viz, aes(x = sex, y= weight, color = sex), size = 2.5, alpha = 0.5) +
  facet_wrap(~grid_full) +
  theme(legend.position = "none") +
  labs(x = "Biological Sex of Juvenile Snowshoe Hares", y = "Weight (grams)", title = "Weight (grams) of Hares by Sex and Site from Bonanza Creek- 1999-2012- JAK") +
  scale_x_discrete(labels = c("Female", "Male", "NA")) +
  geom_point(data = hares_summary, aes(x = sex, y = mean_wt)) +
  geom_errorbar(data = hares_summary,
                aes(x = sex, 
                    ymin = mean_wt-sd_wt, # remove the legend
                    ymax = mean_wt+sd_wt)) + 
  theme_light()
  
                
weightplot + theme(legend.position = "none")

```
Examining the plot of juvenile snowshoe hare weights across the two sexes as well as sites of data collection several trends appear to emerge. First, across all three sites the mean weights for the male hares (indicated by the bold point in the center of the distribution) is greater then both weights recorded for female hares and unidentified individuals (i.e. NA); consistent with our later analyses comparing mean weights across the sexes (mean male weight = mean female weight =. Comparing the three sites, the Black Spruce site recorded the greatest number of hares trapped, Riparian the second greatest, and the Mature site had the least number of trapped hares during the study time frame.



### B. Juvenile weight comparison (male & female snowshoe hares).
**Table 1.** Descriptive Statistics of recored sample weights compared across sex for Juvenile Snowshoe Hares. Data: Kielland et al. (2017) 
```{r, message = FALSE, warning = FALSE, echo =FALSE}
# Creating a lil summary table
hares_summary_final <- juvenile_weight_viz %>% 
  group_by(sex) %>% 
  summarize(mean_wt = mean(weight, na.rm = TRUE), median_wt = median(weight, na.rm = TRUE), sd_wt= sd(weight, na.rm = TRUE),
            sample_size= n())

kable_hare_summ <- hares_summary_final %>% 
  kable(col.names = c("Sex", "Mean", "Median", "Standard Deviation", "N")) %>% 
  kable_styling(full_width = FALSE)

kable_hare_summ


# then I calculate the mean difference
mean_diff <- (948.36 - 865.08)
mean_diff

# Running a t.test on the mean weights and hing length for male and female juvenile hares
mean_weight_male <- hares_juvenile %>% 
  filter(sex == "m") %>% 
  filter(!is.na(weight)) %>% 
  pull(weight)

mean_weight_female <- hares_juvenile %>% 
  filter(sex == "f") %>% 
  filter(!is.na(weight)) %>% 
  pull(weight)

# Running a t.test on the mean weights and hing length for male and female juvenile hares
ttest_mean_weights <- t.test(mean_weight_male, mean_weight_female)
ttest_mean_weights

# Now cohen's d 
cohen_wt <- cohen.d(mean_weight_male, mean_weight_female)
cohen_wt
```
A 2 - 3 sentence comprehensive statement of the means comparison, including at least:
The actual difference in means
Effect size (Cohen's *d*)
Outcome of a two-sample t-test



### C. Relationship between juvenile weight & hind foot length. 
```{r, echo =FALSE, message = FALSE, warning = FALSE, fig.cap= "Figure 3. Chart Comparing the distribution of Juvenile Snowshoe Hares Weights by the length of their Hind feet"}
# Appears to be a positive relationship between weight and hind ft length
cor_weightnhft <- cor.test(hares_juvenile$weight, hares_juvenile$hindft)
cor_weightnhft

# Create a final visualization of juvenile hare hind foot length versus weight. Include a figure caption- UPDATEEEE
plot_weight_hft <- ggplot(hares_juvenile, aes(x = weight, y = hindft, color = sex)) +
  geom_point() +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.2)+
  labs(x = "Hare Weight (grams)", y= "Hind Foot Length (mm)", title = "Relationship between Juvenile Snowshoe Hare Weight and Hind Foot Length- JAK") +
  theme_light()

plot_weight_hft +theme(legend.position = "none")

#Upon seeing your graph, your “boss” tells you to try out linear regression to describe how juvenile snowshoe hare weight changes with respect to hind foot length. Try it out (the only variables you are expected to include here are hind foot length and weight, using data for all juvenile hares in the data, e.g. you are not expected to explore impacts of sex, site, etc.), and check the model diagnostics.
hares_lm <- bonanza_hares %>% 
  select(c("date", "sex", "weight", "hindft", "b_key", "age", "grid", "trap")) %>% 
  filter(age == "j") %>% 
  filter(!is.na(weight)) %>% 
  filter(!is.na(hindft)) %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) 

juvenile_lm <- lm(weight ~ hindft, data = hares_lm)
#summary(juvenile_lm)

juvenile_lm_tidy <- broom::tidy(juvenile_lm)
juvenile_lm_tidy

#juvenile_int <- juvenile_lm_tidy$estimate[1]
#juvenile_int

#juvenile_coef <- juvenile_lm_tidy$estimate[2]

```
We employed simple linear regressions to explore the relationship between mean juvenile hare weight (grams) and mean hind foot length (mm) across all juvenile hares trapped and recorded across all sites. A significant regression model was found (β = 9.52, F(1,247) = 105.2, p < 0.001) with an R2 of 0.2988, meaning that for every 9.52(mm) increase in hind foot length there is a 1 standard deviation increase in hare weight. Assessing the R2 value it appears that hind foot length explains 29% of the variance in hare weight. Though a significant percentage of the variance appears explained, this may indicate that these two variables alone are insufficient to fully explain this relationship. In the future I might suggest adding observations for variables such at least sex and site (if not also age, height, and diet) as we know that these two variables are strongly correlated with mean weight and hindft (while the others I mentioned I would postulate are highly correlated).


## Summary: 
Include a very short summary describing your main findings from your juvenile hares exploration (bulletpoints are fine)


## Citations: 
Data citation: Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

Link to metadata: https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-bnz.55.22

















